---
title: "Covid91 vaccine study Final2021"
author: "Misty O'Brien"
date: "April 27th"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
TrialData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))

TrialSubset =select(TrialData, ! is.numeric)
Alldata= purrr::map(TrialSubset, as.factor)
Alldata = data.frame(Alldata)
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


# Methods
```{r}
maledata= filter(TrialData, sex=="guy")
maledata$sex <- as.factor(maledata$sex)
maledata$infected= factor(maledata$infected)
maledata$infected<- relevel(maledata$infected, "OK")

femaledata= filter(TrialData, sex=="gal")
femaledata$sex <- as.factor(femaledata$sex)
femaledata$infected= factor(femaledata$infected)
femaledata$infected<- relevel(femaledata$infected, "OK")
levels(femaledata$infected)


Druggiedata= filter(TrialData, DrugUser== "yes")
Druggiedata$DrugUsers <- as.factor(Druggiedata$DrugUser)
Druggiedata$infected= factor(Druggiedata$infected)
Druggiedata$infected<- relevel(Druggiedata$infected, "OK")

LGBTQdata= filter(TrialData, LGBTQ== "gay")
LGBTQdata$LGBTQ <- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected= factor(LGBTQdata$infected)
LGBTQdata$infected<- relevel(LGBTQdata$infected, "OK")
```
# Males 

## Bar Chart

```{r}
males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

From this bar chart it shows that out of about 340 males were infected with Covid-19 out of those infected approximately 58% of the males were given the drug, and approximately 42% of males the had gotten the placebo. The males who got the drug were more likely to get Covid-19 than compared to the males who got the placebo. 

### Numerical Summaries  

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

From the table it shows that 340 males who were infected with Covid-19 out of 10,400 males in the study. From the table 3.72% of the males who got the drug contracted Covid-19 and 2.55% of the males who go the placebo contracted Covid-19. It appears that was not an effective drug, due to the table showing that males who got the drug were about 50% more likely to contract Covid-19. This was not an effective drug for males.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi squared table shows that the p-value is 0.0006181 is very small that is it very unlikely that the result are by chance. 
$H_0$ The null hypothesis is that taking the drug, there is no difference of males contracting Covid-19 or not.
$H_A$ Th alternative hypothesis is that the drug in fact does make a difference a male contracts Covid-19
Therefore, we reject the null hypothesis 

The odd ratio by the Fishers Exact Test is 1.47:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formala O/(1+O)x 100 you get 59.5% effectiveness of the Covid91 drug

Ultimately the drug is 40% effective in the male population, which is completely ineffective

# Females 

## Bar Chart

```{r}
females <- femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of approximately 100 females who were infected with Covid-19, 33% were given the drug and 67% were given the placebo.


### Numerical Summaries  

```{r}
table2 <- xtabs(~infected + treatment, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```
From the table it shows that 607 females who were infected with Covid-19 out of 9,213 females in the study. 4.44% of the females who got the drug contracted Covid-19 and 8.67% of the females who go the placebo contracted Covid-19. It appears that was  an effective drug, due to the table showing that females who got the drug were about 50% less likely to contract Covid-19. This is an effective drug for females.


## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The p-value is small at 0.0000000000000002 that it is extremely unlikely the result were to happen by chance.

$H_0$ The null hypothesis is that the drug makes no difference in the females population contracting Covid-19 or not.

$H_A$ Th alternative hypothesis is that the drug in fact does make a difference in whether or not a female contracts Covid-19.
Therefore, we null hypothesis

The fisher exact test odd ratio is 2.05:1, when taking the odd ratio and computing it into a percent of effectiveness by using the formula O/(1+O)x 100 you get 67% of effective for females.
Ultimately the drug is 67% effective in the female population, which is just a little bit effective.

# LGBTQ

## Bar Chart

```{r}
gay <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(gay,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(gay,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of approximately 100 LGBTQ who were infected with Covid-19, 40% were given the drug and 30% were given the placebo.


### Numerical Summaries  

```{r}
table3 <- xtabs(~treatment + infected, data=LGBTQdata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

From the table it shows that 116 LGBTQ who were infected with Covid-19 out of 2,040 LGBTQ in the study. From the table 70.6% of the LGBTQ who got the drug contracted Covid-19 and 29.3% of the LGBTQ who go the placebo contracted Covid-19. It appears that was not an effective drug, due to the table showing that LGBTQ who got the drug were about 50% more likely to contract Covid-19. This was not an effective drug for LGBTQ population.

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```
The p-value is small at 0.0000057 that it is extremely unlikely the result were to happen by chance.

$H_0$ The null hypothesis is that the drug makes no difference in the LGBTQ population contracting Covid-19 or not.

$H_A$ Th alternative hypothesis is that the drug in fact does make a difference in whether or not a LGBTQ Covid-19.
Therefore, we null hypothesis

The fisher exact test odd ratio is 0.40:1, when taking the odd ratio and computing it into a percent of effectiveness by using the formula O/(1+O)x 100 you get 29% of effective for LGBTQ.
Ultimately the drug is 29% effective in the LGBTQ population, which is just completely ineffective.

# DrugUsers

## Bar Chart

```{r}
druggies <- Druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

From this bar chart it shows that out of about 100 drug users were infected with Covid-19 out of those infected approximately 32% of the drug users were given the vaccine, and approximately 68% of drug users the had gotten the placebo. The drug users who got the vaccine were more likely to get Covid-19 than compared to the drug users who got the placebo. 


### Numerical Summaries  

```{r}
table4 <- xtabs(~treatment + infected, data=Druggiedata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

From the table it shows that 117 drug users who were infected with Covid-19 out of 2,283 drug users in the study. 33.33% of the drug users who got the vaccine contracted Covid-19 and 66.67% of the drug users who go the placebo contracted Covid-19. It appears that was an effective vaccine, due to the table showing that drug users who got the vaccine were about 67% less likely to contract Covid-19. This was an effective vaccine for drug users.

## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

 The p-value is small at 0.002273 that it is extremely unlikely the result were to happen by chance.

$H_0$ The null hypothesis is that the drug/vaccine makes no difference in the DrugUsers population contracting Covid-19 or not.

$H_A$ Th alternative hypothesis is that the drug/vaccine in fact does make a difference in whether or not a DrugUsers contracts Covid-19.
Therefore, we null hypothesis
The fisher exact test odd ratio is 1.83:1, when taking the odd ratio and computing it into a percent of effectiveness by using the formula O/(1+O)x 100 you get 60% of effective for females.
Ultimately the vaccine is 60% effective in the DrugUsers population, which is just a little bit effective.
 
# Discussion and Conclusion

When looking at four population and found that the only females and DrugUsers population had high effectiveness of the vaccine. Females had an effectiveness at 67% and Drugusers effectiveness at 60%. Both the male population and LGBTQ population showed that the vaccine was not beneficial for them and and were ineffective in protecting them from contracting Covid-19. The males had a 40% of effectiveness, while LGBTQ had a 29% of effectiveness, which is not all effective.


